---
title: "Problem_Set_4"
author: "Zongyi Liu"
date: "2023-10-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1: Tidyverse

```{r}
library(nycflights13)
library(tidyverse)
```

## a. Generate Tables

Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport.

```{r}
table_1 <-
  flights %>%
  group_by(dest) %>%
  summarise(Mean = mean(dep_delay, na.rm = TRUE),Median = median(dep_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c(dest = "faa"))%>%
  select(Mean,Median,name)%>%
  arrange(desc(Mean))

print(table_1, n = nrow(table_1))
```

Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport. Exclude any destination with under 10 flights. Do this exclusion through code, not manually.

```{r}
table_2 <-
  flights %>%
  group_by(dest) %>%
  # arrival delay NA's are cancelled flights
  summarise(Mean = mean(arr_delay, na.rm = TRUE),Median = median(arr_delay, na.rm = TRUE),totalflights=n())%>%filter(totalflights>9)%>%
  inner_join(airports, by = c(dest = "faa"))%>%
  select(Mean,Median,name)%>%
  arrange(desc(Mean))
print(table_2, n = nrow(table_2))
```

Also,

-   Order both tables in descending mean delay.

-   Both tables should use the airport *names* not the airport *codes*.

-   Both tables should print all rows.

## b

How many flights did the aircraft model with the fastest average speed take? Produce a tibble with 1 row, and entries for the model, average speed (in MPH) and number of flights.

```{r}
library(nycflights13)
library(dplyr)
```

```{r}
flights %>%left_join(planes, by = "tailnum")%>%group_by(model) %>% select(speed, tailnum)%>%summarize(num_flights = n())%>%arrange (desc(num_flights))%>%as_tibble()
```

## Problem 2: `get_temp()`

Load the Chicago NNMAPS data we used in the visualization lectures. Write a function `get_temp()` that allows a user to request the average temperature for a given month. The arguments should be:

-   `month`: Month, either a numeric 1-12 or a string.

-   `year`: A numeric year.

-   `data`: The data set to obtain data from.

-   `celsius`: Logically indicating whether the results should be in celsius. Default `FALSE`.

-   `average_fn`: A function with which to compute the mean. Default is `mean`.

The output should be a numeric vector of length 1. The code inside the function should, as with the rest of this problem, use the **tidyverse**. Be sure to sanitize the input.

```{r}
library(dlnm)
data(chicagoNMMAPS)

nnmaps<-chicagoNMMAPS
```

```{r}
library(tidyverse)

# Define a function to get the average temperature
get_temp <- function(month, year, data, celsius = FALSE, average_fn = mean) {
  # Sanitize input
  if (!is.numeric(year) || year <= 0) {
    stop("Invalid year. Please provide a valid numeric year.")
  }
  
  # Convert month to numeric if it's a string
  if (is.character(month)) {
    month <- match(tolower(month), tolower(month.abb))
    if (is.na(month)) {
      stop("Invalid month name. Please provide a valid month.")
    }
  }
  
  # Filter the data for the specified month and year
  filtered_data <- data %>%
    filter(month == month, year == year) #  
  
  if (nrow(filtered_data) == 0) {
    stop("No data available for the specified month and year.")
  }
  
  # Calculate the average temperature using the provided function
  avg_temp <- filtered_data %>%
    summarise(average_temp = average_fn(Temperature)) %>%
    pull(average_temp)
  
  # Convert to Celsius if requested
  if (celsius) {
    avg_temp <- (avg_temp - 32) * 5/9
  }
  
  return(avg_temp)
}

# Usage example:
# Get the average temperature for April 1999 in Fahrenheit
avg_temp_fahrenheit <- get_temp(month = "Apr", year = 1999, data = nnmaps)
print(avg_temp_fahrenheit)

# Get the average temperature for April 1999 in Celsius
avg_temp_celsius <- get_temp(month = "Apr", year = 1999, data = nnmaps, celsius = TRUE)
print(avg_temp_celsius)
```

# Problem 3 SAS

This problem should be done entirely within SAS.

Access the [RECS 2020 data](https://www.eia.gov/consumption/residential/data/2020/index.php?view=microdata) and download a copy of the data. You may import the CSV or load in the `sas7bdat` file directly. (This is **not** the 2009 version we used in lecture.) You'll probably also need the "Variable and response cookbook" to identify the proper variables. Load or import the data into SAS.

a.  What state has the highest percentage of records? What percentage of all records correspond to Michigan? (Don't forget to account for the sampling weights!)

b.  Generate a histogram of the total electricity cost in dollars, amongst those with a strictly positive cost.

c.  Generate a histogram of the log of the total electricity cost.

d.  Fit a linear regression model predicting the log of the total electricity cost based upon the number of rooms in the house and whether or not the house has a garage. (Don't forget weights.)

e.  Use that model to generate predicted values and create a scatterplot of predicted total electricity cost vs actual total electricity cost (**not** on the log scale).

```{r}
data<-read.csv("recs2020_public_v5.csv")
data
```

# Problem 4 Multiple Tools

Take a look at the Codebook. For very minor extra credit, how was the Codebook generated? (No loss of points if you skip this.)

> The codebook contains the variable names, labels, and tabulations of responses for the questions asked in the survey.
