---
title: "Problem_Set_5"
author: "Zongyi Liu"
date: "2023-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1 Plotting

Use the `nnmaps` data set again. I'd recommend using `ggplot2` for these.

### Question a

Produce a graph of the `mean` monthly temperature in celsius. Draw a scatter-plot where the x-axis is month and the y-axis is the average monthly temperature in celsius. Add a line connecting the points within each season and color the lines and points by season (same color for lines and points in the same season). Be sure both x-axis and the legend are ordered appropriately, and that all labels/titles are clear and ready for publication (e.g. no variable names).

```{r}
data_use<-read.csv("chicago-nmmaps.csv")
#data_use
library(ggplot2)
library(dplyr)

data_use_2<-data_use%>%mutate(
  temp_cel=(temp-32)*5/9,
  date = as.POSIXct(date, format = "%Y-%m-%d"),
  year_month = format(date, "%Y-%m")
)%>%group_by(month,season)

# Reorder the month column
# data_use_2$month%>%reorder(order = month.abb)

data_use_2$month <- factor(data_use_2$month, levels = month.abb)

data_use_3<-data_use_2%>%summarise(avg_temp_cel=mean(temp_cel))
```

```{r}
ggplot(data_use_3, aes(x = month, y = avg_temp_cel, color = season, group = season)) +
    geom_point() +
    geom_line() +
    labs(
        x = "Time",
        y = "Average Temperature (C)",
        title = "Average Monthly Temperature in Chicago",
        color = "Season"
    )
```

### Question b

Produce a similar plot with four sets of lines: mean monthly temperature in celsius, mean monthly O3, mean monthly PM10, and mean monthly dewpoint. Figure out a way to distinguish between the four lines. Again, make sure the plot is publication-ready (clean and clear labels). Use the plot to answer the following question: "Which of the four variables seems to have the least seasonal trend?"

```{r}
data_use_4<-data_use_2%>%
  group_by(month,season)%>%
  summarise(
    avg_temp_cel=mean(temp_cel,na.rm = T),
    avg_O3 = mean(o3, na.rm =T),
    avg_PM10 = mean(pm10, na.rm =T),
    avg_dewpoint = mean(dewpoint, na.rm = T)
    )

data_use_4
```

```{r}
# Totally the same as before
ggplot(data_use_4) + 
    geom_point(aes(x = month, y = avg_temp_cel, color = season)) +
    geom_line(aes(x = month, y = avg_temp_cel, color = season, group = season, linetype = "Temperature")) +
    geom_point(aes(x = month, y = avg_O3, color = season)) +
    geom_line(aes(x = month, y = avg_O3, color = season, group = season, linetype = "O3")) +
    geom_point(aes(x = month, y = avg_PM10, color = season)) +
    geom_line(aes(x = month, y = avg_PM10, color = season, group = season, linetype = "PM10")) +
    geom_point(aes(x = month, y = avg_dewpoint, color = season)) +
    geom_line(aes(x = month, y = avg_dewpoint, color = season, group = season, linetype = "Dew_Point")) +
    labs(
        x = "Month",
        y = "Value",
        title = "Monthly Average Weather Data in Chicago",
        color = "Season",
        linetype = "Weather Data"
    )
```

> From the plot we can say that O3 has the least seasonal trend

# Problem 3 NYCFlights 13

### Question a

Use the **tidyverse** for this problem. In particular, use piping and **dplyr** as much as you are able. **Note**: Use of any deprecated functions will result in a point loss.

Install and load the package [**nycflights13**](https://cran.r-project.org/package=nycflights13).

Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport. Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport. Exclude any destination with under 10 flights. Do this exclusion through code, not manually.

Additionally,

-   Order both tables in descending mean delay.

-   Both tables should use the airport *names* not the airport *codes*.

-   Both tables should print all rows.

```{r}
library(nycflights13)
library(data.table)
```

```{r}
# Convert flights and airports to data.table
flights_dt <- as.data.table(flights)
airports_dt <- as.data.table(airports)

# Filter out airports with less than 10 flights
valid_origins <- flights_dt[, .N, by = .(origin)][N >= 10, origin]
filtered_flights <- flights_dt[origin %in% valid_origins]

dep_delay_table <- filtered_flights[, .(mean_dep_delay = mean(dep_delay, na.rm = TRUE),
                                        median_dep_delay = median(dep_delay, na.rm = TRUE)
                                    ), by = .(origin)]
# Add airport names
dep_delay_table <- airports_dt[dep_delay_table, on = .(faa = origin), 
                                .(name, mean_dep_delay, median_dep_delay)]
# Order by mean departure delay
dep_delay_table <- dep_delay_table[order(-mean_dep_delay)]
dep_delay_table
```

### Question b

How many flights did the aircraft model with the fastest average speed take? Produce a tibble with 1 row, and entires for the model, average speed (in MPH) and number of flights.
